MEPS

### 2022-11-29 scripts/00_eda.qmd
Implemented suppression flowchart

```{r}
# Libraries
library(tidyverse)
library(haven)

# Read in data
# For the time being, just going to work with 2001 - 2006 to recreate table
# in that paper
dat_files <- list.files(path = 'data', pattern = 'meps', full.names = TRUE)
dat_names <- map_chr(dat_files, ~str_extract(.x, 'meps[0-9]{4}'))
# 
readAndSelect <- function(data_path, data_year) {
  yr <- substr(data_year, 7, 8)
  eye_var <- paste0('DSEY', yr, '53')
  chol_var <- paste0('DSCH', yr, '53')
  feet_var <- paste0('DSFT', yr, '53')
  flu_var <- paste0('DSFL', yr, '53')
  if (str_ends(data_path, '[sSpP]{3}') == TRUE) {
    temp <- read_xpt(data_path)

  } else {
    temp <- read_sas(data_path)

  }
  if (yr == '01') {
    temp <- temp %>% 
      mutate(RACEX = recode_factor(
        RACEX,
        `1` = 'American Indian',
        `2` = 'Aleut, Eskimo',
        `3` = 'Asian or Pacific Islander',
        `4` = 'Black',
        `5` = 'White'
      ))
  }

  temp %>% 
    mutate(across(starts_with('RACE'), ~as.factor(.x))) %>% 
    select(starts_with('DOB'),
           starts_with('DUPERSID'),
           starts_with('USBORN'),
           starts_with('BORN'), # Two different born in USA variables used in different years
           starts_with('AGE'),
           starts_with('RACE'),
           starts_with('SEX'),
           starts_with('VARPSU'), # PSU
           starts_with('VARSTR'), # Strata
          # starts_with('INSCOV'), # Insurance coverage variable
           starts_with('HIDEG'), # Highest degree - want only HIDEGYR and HIDEG
           starts_with('HIDEG'), # Highest degree when first entered MEPS
           starts_with('TTLP'), # Total income
           starts_with('POVCAT'), # Categorical poverty status
           #starts_with('POVLEV'), # Continuous poverty status
           #starts_with('REGION'),
           starts_with('DENTCK53'), # Dental checkup, only 2000 - 2016,
           matches('DVTOT[0-9]{2}'), # # of dental care visits in YYYY        
           starts_with('CHOLCK'), # Cholesterol check, only 2000 - 2016,
           starts_with('CHECK53'), # Routine check up, only 2000 - 2016,
           starts_with('FLUSHT'), # Flu shot. only 2000 - 2016
           starts_with('DSFL'),
           starts_with('DIAB'), # 3 different variable names for diabetes, all start with DIABDX. this should also grab the DAIBW weight variables
           starts_with(!!eye_var), # Get only the eye variable for each year
           starts_with(!!flu_var),
           starts_with(!!chol_var),
           starts_with(!!feet_var),
           starts_with('DSEYE53'), # Have to get the 2001 eye check variable too
           starts_with('DSA1'), # A1C test,
           starts_with('DSCKFT'), # Feet check, only 2001 - 2007
          -starts_with('DSEB'), # Remove the eye test in year before columns
          -contains('AGE31X'),
          -contains('AGE42X'),
          -contains('AGE53x'), # Remove some of the extra age variables
           starts_with('PERWT'), # Person weight level. Not sure I need this, since just working with DCS variables
           starts_with('INSCOP') # Inscope, again not sure if needed but putting here anyway for now
           )
}
safeReadAndSelect <- safely(readAndSelect)
dat_list <- map2(dat_files, dat_names, ~readAndSelect(.x, .y))
names(dat_list) <- dat_names

# Check number of columns
map(dat_list, ~ncol(.x))

```
# To do for 11/4 or 11/7: Figure out how to actually recode DSEYE53.

```{r}
dat <- bind_rows(dat_list, .id = 'year')
# Do some type conversions
dat <- dat %>%
  mutate(
    DSEY0153 = case_when(
      DSEYE53 == 2 ~ 1,
      DSEYE53 %in% c(1, 3, 4, 5) ~ 2,
      DSEYE53 == -9 ~ -9,
      DSEYE53 == -8 ~ -8,
      DSEYE53 == -1 ~ -1,
      TRUE ~ as.numeric(DSEYE53)
    ),
    DSA1C53 = as.factor(
      case_when(
        DSA1C53 == -1 ~ 'Inapplicable',
        DSA1C53 == -7 ~ 'Refused',
        DSA1C53 == -8 ~ 'DK',
        DSA1C53 == -9 ~ 'Not ascertained',
        DSA1C53 == -15 ~ 'Cannot be computed',
        DSA1C53 %in% c(0, 1, 96) ~ 'Less than 2 or more A1C tests in a year',
        DSA1C53 >= 2 ~ '2 or more A1C tests in a year',
        TRUE ~ as.character(DSA1C53)
      )
    ),

    DSCKFT53 = as.factor(
      case_when(
        DSCKFT53 >= 1 ~ 'Within last year',
        DSCKFT53 == 0 ~ 'No foot examinations in a year',
        DSCKFT53 == -1 ~ 'Inapplicable',
        DSCKFT53 == -7 ~ 'Refused',
        DSCKFT53 == -8 ~ 'DK',
        DSCKFT53 == -9 ~ 'Not ascertained'
      )
    )
  ) %>%
  select(-DSEYE53) %>% # Remove the 2001 eye variable that we recoded
  mutate(
    across(
      c(
        starts_with('DS'),
        CHOLCK53,
        CHECK53,
        FLUSHT53,
        HIDEGYR,
        HIDEG,
        RACEX,
        RACETHNX,
        USBORN42
      ),
      ~ as.factor(.x)
    )
  )

# Recode factors to facilitate working with them
dat <- dat %>%
  mutate(
    across(
      intersect(starts_with('DSEY'),
                ends_with('53')),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-1` = 'Inapplicable',
        `1` = 'YES',
        `2` = 'NO'
      )
    ),
    across(
      intersect(starts_with('DSCH'),
                ends_with('53')),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-1` = 'Inapplicable',
        `1` = 'Within last year',
        `2` = 'NO'
      )
    ),
    across(
      intersect(starts_with('DSFL'),
                ends_with('53')),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-1` = 'Inapplicable',
        `1` = 'Within last year',
        `2` = 'NO'
      )
    ),
    across(
      intersect(starts_with('DSFT'),
                ends_with('53')),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-1` = 'Inapplicable',
        `1` = 'Within last year',
        `2` = 'NO'
      )
    ), 
    across(
      starts_with('DVTOT'),
      ~ as.factor(case_when(.x >= 1 ~ 1,
                            TRUE ~ 0)) %>%
      recode_factor(`1` = 'Once a year',
                    `0` = 'Less than once a year')
    ), 
    across(
      starts_with('HIDE'),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-7` = 'Refused',
        `1` = 'No Degree',
        `2` = 'GED',
        `3` = 'HS',
        `4` = 'Bachelor',
        `5` = 'Masters',
        `6` = 'Doctorate',
        `7` = 'Other',
        `8` = 'Under 16, inapplicable'
      ) %>%
        fct_collapse(
          `Less than high school` = c('No Degree', 'GED', 'Under 16, inapplicable'),
          `High school` = c('HS'),
          `Greater than HS` = c('Bachelor', 'Masters', 'Doctorate', 'Other'),
          `Not available` = c('Not ascertained', 'DK', 'Refused', '-1', '-15')
        )
    ),
    across(
      c(RACEX, RACEV1X),
      ~ recode_factor(
        .x,
        `1` = 'White',
        `2` = 'Black',
        `3` = 'Amer Indian/Alaska Native',
        `4` = 'Asian',
        `5` = 'Native Hawiian/Pacific Islander',
        `6` = 'Multiple races reported',
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-7` = 'Refused'
      )
    ),
    RACETHNX = recode_factor(
      RACETHNX,
      `1` = 'Hispanic',
      `2` = 'Black/Not Hispanic',
      `3` = 'Asian/Not Hispanic',
      `4` = 'Other Race/Not Hispanic',
    ) %>% fct_expand('White/Not Hispanic'),
    RACETHX = recode_factor(
      RACETHX,
      `1` = 'Hispanic',
      `2` = 'White/Not Hispanic',
      `3` = 'Black/Not Hispanic',
      `4` = 'Asian/Not Hispanic',
      `5` = 'Other Race/Not Hispanic',
    ),
    USBORN42 = recode_factor(
      USBORN42,
      `-9` = 'Not ascertained',
      `-8` = 'DK',
      `-7` = 'Refused',
      `-1` = 'Inapplicable',
      `1` = 'Yes',
      `2` = 'No'
    ),
    BORNUSA = recode_factor(
      BORNUSA,
      `-15` = 'Cannot be computed',
      `-9` = 'Not ascertained',
      `-8` = 'DK',
      `-7` = 'Refused',
      `-1` = 'Inapplicable',
      `1` = 'Yes',
      `2` = 'No'
    ),
    across(
      c(CHOLCK53, CHECK53, FLUSHT53),
      ~ recode_factor(
        .x,
        `-9` = 'Not ascertained',
        `-8` = 'DK',
        `-7` = 'Refused',
        `-1` = 'Inapplicable',
        `1` = 'Within last year',
        `2` = 'Within past 2 years',
        `3` = 'Within past 3 years',
        `4` = 'Within past 5 years',
        `5` = 'More than 5 years',
        `6` = 'Never'
      ) %>%
        fct_collapse(
          `Within last 2 more more years` = c(
            'Within past 2 years',
            'Within past 3 years',
            'Within past 5 years',
            'More than 5 years',
            'Never'
          ),
          `Not available` = c(
            'Not ascertained',
            'DK',
            'Refused',
            'Inapplicable'
          )
        )
    ),
    DENTCK53 = recode_factor(
      DENTCK53,
      `-9` = 'Not ascertained',
      `-8` = 'DK',
      `-7` = 'Refused',
      `-1` = 'Inapplicable',
      `1` = 'Twice a year or more',
      `2` = 'Once a year',
      `3` = 'Less than once a year',
      `4` = 'Never go to dentist'
    ),
    SEX = recode_factor(SEX,
                        `1` = 'Male',
                        `2` = 'Female')

  )

dat$RACETHNX[dat$RACEX == 'White'] <- 'White/Not Hispanic'

# Need to coalesce the eye exam columns into a single column for a Y/N eye exam

@@

@@

```{r}
# Three columns denote the DIAB diagnosis based on different years. I coalesce them
# to create a single column, which we can then filter on.
# Coding appears the same for all variables. 1 is diagnosed with diabetes, 2 is not. Other numbers
# are variations on missing or non-response
dat$DIABDX_all <- coalesce(dat$DIABDX53, dat$DIABDX_M18) # When using full data, add dat$DIABDX_M18,

# Variable to coallesce
# HIDEGYR/HIDEG
# POVCAT
# DIABW variables
dat <- dat %>%
  mutate(
    DSEY_all = coalesce(!!!select(., starts_with('DSEY'))),
    HIDEG_all = coalesce(!!!select(., starts_with('HIDEG'))),
    DIABW = coalesce(!!!select(., starts_with('DIABW'))),
    POVCAT_all = coalesce(!!!select(., starts_with('POVCAT'))),
    USBORN_all = coalesce(!!!select(., contains('BORN'))),
    VARPSU_all = coalesce(!!!select(., starts_with('VARPSU'))),
    VARSTR_all = coalesce(!!!select(., starts_with('VARSTR'))),
    AGE_all = coalesce(!!!select(., starts_with('AGE'))),
    RACE_all = coalesce(!!!select(., RACETHNX, RACETHX)),
    CHOL_all = coalesce(!!!select(., CHOLCK53, starts_with('DSCH'))),
    FLU_all = coalesce(!!!select(., FLUSHT53, starts_with('DSFL'))),
    DENT_all = coalesce(!!!select(., DENTCK53, num_range('DVTOT', 17:20))),
    FEET_all = coalesce(!!!select(., DSCKFT53, starts_with('DSFT'))),

  ) %>% 
  mutate(
    AGE_all = as.factor(case_when(
      AGE_all == -1 ~ NA_character_,
      AGE_all < 18 & AGE_all > -1 ~ NA_character_,
      AGE_all >= 18 & AGE_all <= 44 ~ '18 to 44',
      AGE_all >= 45 & AGE_all <= 64 ~ '45 to 64',
      AGE_all >= 65 & AGE_all <= 74 ~ '65 to 74',
      AGE_all >= 75 ~ '75+',
      TRUE ~ as.character(AGE_all)
    ))
  )
# Get just data for people with diabetes
# Get just the variables I want
diab_dat <- dat %>% 
  select(year, DUPERSID, SEX, DIABW, ends_with('_all'),
         CHOLCK53, FLUSHT53, CHECK53, DENTCK53, DSA1C53, DSCKFT53)

@@

```{r}
library(sitrep)
library(gtsummary)
library(srvyr)

diab_list <- group_split(diab_dat, year)
diab_svy <- diab_list[[17]] %>%
  as_survey_design(ids = VARPSU_all, weights = DIABW, strata = VARSTR_all, nest = TRUE)
diab_only <- subset(diab_svy, DIABW != 0)
#svytable(~SEX, design = diab_only)
#test <-svytable(~CHOLCK53 + USBORN_all, design = diab_only, Ntotal = 100)

test <- diab_svy %>%
  tab_survey(
    #CHOL_all,
    #FLU_all,
    #DENT_all,
    #DSA1C53,
    DSEY_all,
    #FEET_all,
    pretty = FALSE,
    strata = 'AGE_all',
    wide = FALSE,
    method = 'mean',
    #drop = c('Not ascertained', 'DK', 'Refused', 'Inapplicable'),
    keep = c(
      'Within last year',
      '2 or more A1C tests in a year',
      '1 or more foot examinations in a year',
      'Once a year',
      #Dental checkups
      'YES' # eye exam
    )
  )
```

@@

# Get the dashboard medians
```{r}
dash_files <- list.files(here::here('data', 'dashboard'), full.names = TRUE)
dash_file_names <- str_extract(list.files(here::here('data', 'dashboard')), '^([^.])+')
dash_list <-
  map(dash_files,
      ~ read_csv(
        .x,
        skip = 2,
        col_names = TRUE,
        n_max = 21,
        col_types = c('dcddd'),
        na = c('Suppressed', 'No Data')
      ))
names(dash_list) <- dash_file_names
dash_dat <- bind_rows(dash_list, .id = 'file_name')%>% 
  separate(file_name, into = c('indicator', 'stratifier', 'strata'), sep = "_") %>% 
  janitor::clean_names() %>% 
  mutate(across(percentage:upper_limit, ~.x/100))

@@

>>>>

```{r}
# Libraries
library(tidyverse)
library(haven)
library(targets)

tar_load(dat_list)

@@

# To do for 11/4 or 11/7: Figure out how to actually recode DSEYE53.

```{r}
tar_load(cleaned_dat)
```

@@

```{r}
tar_load(diab_dat)

ggplot(dplyr::filter(sex_indicator, variable == 'eye-exam', value %in% c(
    'Within last year',
    '2 or more A1C tests in a year',
    '1 or more foot examinations in a year',
    'Once a year',
    #Dental checkups
    'YES'  ))) + geom_line(aes(x = year, y = proportion, group = SEX))

@@

```{r}
library(survey)
library(sitrep)
library(gtsummary)
library(srvyr)

diab_list <- group_split(diab_dat, year)
diab_all_svy <- diab_dat %>% 
  as_survey_design(ids = VARPSU_all, weights = DIABW, strata = VARSTR_all, nest = TRUE)
diab_svy <- diab_list[[17]] %>%
  as_survey_design(ids = VARPSU_all, weights = DIABW, strata = VARSTR_all, nest = TRUE)
diab_only <- subset(diab_svy, DIABW != 0)
#svytable(~SEX, design = diab_only)
#test <-svytable(~CHOLCK53 + USBORN_all, design = diab_only, Ntotal = 100)

tbl_svysummary(
  diab_all_srvy,
  by = year,
  include = c(#CHOL_all,
              # FLU_all,
              # DENT_all,
              DSA1C53
              # DSEY_all,
              # FEET_all
              ),
  statistic = list(all_categorical() ~ "{n}")
)
tbl_svysummary(diab_svy)

test <- diab_svy %>%
  tab_survey(
    #CHOL_all,
    #FLU_all,
    #DENT_all,
    #DSA1C53,
    DSEY_all,
    #FEET_all,
    pretty = FALSE,
    strata = 'AGE_all',
    wide = FALSE,
    statistic = list(DSEY_all ~ "{n}"),
    method = 'mean',
    #drop = c('Not ascertained', 'DK', 'Refused', 'Inapplicable'),
    keep = c(
      'Within last year',
      '2 or more A1C tests in a year',
      '1 or more foot examinations in a year',
      'Once a year',
      #Dental checkups
      'YES' # eye exam
    )
  )
```

@@

# Get the dashboard medians

```{r}
dash_files <- list.files(here::here('data', 'dashboard'), full.names = TRUE)
dash_file_names <- str_extract(list.files(here::here('data', 'dashboard')), '^([^.])+')
dash_list <-
  map(dash_files,
      ~ read_csv(
        .x,
        skip = 2,
        col_names = TRUE,
        n_max = 21,
        col_types = c('dcddd'),
        na = c('Suppressed', 'No Data')
      ))
names(dash_list) <- dash_file_names
dash_dat <- bind_rows(dash_list, .id = 'file_name')%>% 
  separate(file_name, into = c('indicator', 'stratifier', 'strata'), sep = "_") %>% 
  janitor::clean_names() %>% 
  mutate(across(percentage:upper_limit, ~.x/100))

@@

```{r}
